01_pca

### 2019-05-25 analysis/02_pca_all.Rmd
not quite

theme_rmh <- function(){ 
    theme_bw(base_size=14) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "right",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

treatmentcolors <-  c("control" = "#bdbdbd", 
                              "bldg" =  "#d95f02", 
                              "lay"  = "#a1d99b", 
                              "inc.d3" = "#74c476", 
                              "inc.d9"  = "#41ab5d", 
                              "inc.d17" = "#238b45", 
                              "hatch"  = "#807dba", 
                              "n5"  = "#6a51a3", 
                               "n9" = "#4a1486")

@@

```{r rawdata}
counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)

samples$treatment <- factor(samples$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))
samples$tissue <- factor(samples$tissue, levels = 
                              c("hypothalamus", "pituitary", "gonad"))

geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)
```

@@

## subsets of normalized data show by tissue and sex

```{r tissuesex}
plotlist = list()
for (eachgroup in levels(samples$treatment)){

  print(eachgroup)

  colData <- samples %>%
      dplyr::filter(treatment == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  countData <- pseudocounts %>% dplyr::select(one_of(savecols))

  p <- autoplot(pca, data = colData, alpha = 'tissue', shape = "sex", color = "treatment")

en <- p + labs(subtitle = eachgroup) +
  theme_rmh() +
    theme_rmh()  +
  stat_ellipse(aes(linetype = sex, alpha = tissue, color = treatment)) +
   theme_bw(base_size=8) +
    theme(legend.position = "bottom") +
  scale_color_manual(values = treatmentcolors) + 
  guides(color = FALSE)

plotlist[[eachgroup]] = en

}

@@

  countData <- pseudocounts %>% dplyr::select(one_of(savecols))

@@

  p <- autoplot(pca, data = colData, colour = 'treatment', shape = "sex")

en <- p + labs(subtitle = eachtissue) +
  theme_rmh() +  guides(colour = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = treatment)) +
  theme_bw(base_size=8) +
    theme(legend.position = "bottom") +
  scale_color_manual(values =  treatmentcolors)

>>>>

source("../R/themes.R")  # load custom themes and color palletes

@@

```{r rawdata}
samples <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
counts <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1, header = T)

samples$treatment <- factor(samples$treatment, levels = 
                             c("control", "bldg", "lay",
                               "inc.d3", "inc.d9", "inc.d17", 
                               "hatch", "n5",  "n9",
                               "m.inc.d3", "m.inc.d8","m.inc.d9",
                               "m.inc.d17","m.n2" ,"prolong", "extend"))
samples$tissue <- factor(samples$tissue, levels = 
                             c("hypothalamus", "pituitary", "gonad" ))

samples$lastday <- ifelse(grepl("m.inc.d3|m.inc.d9|m.inc.d17|m.n2", samples$treatment), "empty nest", 
                    ifelse(grepl("m.inc.d8|hatch|extend", samples$treatment),"chicks hatch",      
                     ifelse(grepl("n5", samples$treatment),"chicks early",
                     ifelse(grepl("n9", samples$treatment),"chicks later",
                     ifelse(grepl("control", samples$treatment),"control",
                     ifelse(grepl("bldg", samples$treatment),"nest building",
                     ifelse(grepl("prolong", samples$treatment),"eggs delay",
                      ifelse(grepl("lay", samples$treatment),"eggs lay",
                      ifelse(grepl("inc.d3", samples$treatment),"eggs early",
                     ifelse(grepl("inc.d9", samples$treatment),"eggs middle",
                    ifelse(grepl("inc.d17", samples$treatment),"eggs later", NA)))))))))))

samples$penultimate <-  ifelse(grepl("extend", samples$treatment),"eggs delay",  
                          ifelse(grepl("m.n2", samples$treatment),"chicks hatch",
                     ifelse(grepl("n5", samples$treatment),"chicks early",
                     ifelse(grepl("n9", samples$treatment),"chicks later",
                     ifelse(grepl("control", samples$treatment),"control",
                     ifelse(grepl("bldg|lay", samples$treatment),"nest building",
                     ifelse(grepl("prolong", samples$treatment),"eggs delay",
                      ifelse(grepl("inc.d3|m.inc.d3", samples$treatment),"eggs early",
                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", samples$treatment),"eggs middle",
                    ifelse(grepl("inc.d17|hatch", samples$treatment),"eggs later", NA))))))))))

samples$xlabel <- samples$treatment

levels(samples$xlabel) <-  c("control", "nest.building", "egg.lay",
                               "eggs.early", "eggs.early.remove", 
                               "eggs.mid", "eggs.mid.hatch", "eggs.mid.remove",
                               "eggs.end", "eggs.end.remove",
                               "chicks.hatch",  "chicks.hatch.remove"  ,
                               "chicks.mid", "eggs.delay", "eggs.delay.hatch", "chicks.end")

samples$lastday <- factor(samples$lastday, levels =  c("control", "nest building", "eggs lay",
                               "eggs early", "eggs middle",  "eggs later","eggs delay",
                               "chicks hatch", "chicks early",  "chicks later", "empty nest"))

samples$penultimate <- factor(samples$penultimate, levels =  c("control", "nest building",
                               "eggs early",  "eggs middle",  "eggs later",  "eggs delay",
                               "chicks hatch", "chicks early",  "chicks later"))

@@

@@

  countData <- counts %>% dplyr::select(one_of(savecols))

@@

  p <- autoplot(pca, data = colData, fill = lastday, color = penultimate, shape = "tissue")

en <- p + labs(subtitle = eachtissue) +
  theme_rmh() +  guides(color = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  theme_bw(base_size=8) +
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = colorlastday) +
  scale_color_manual(values = colorpenultimate)
